Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2022-12-05 Manuscript_rev1.tex
Add files via upload

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
	\caption{Example dimer in orientation $\bar{\bf s}$}
	\label{fig:dimer}
\end{figure}

With any trapped particle there are two main sources of motion,
Brownian and optical forces.  The dimer's Brownian motion
($\Delta q^{\textbf{B}}$) is simulated by generating a set of
normally-distributed numbers whose covariance is given by the dimer's
diffusion tensor $\textbf{D}$.  The diffusion tensor is computed based
on Farsund's work \cite{Farsund1996}. For computing the optical force
Brownian OT uses the T-Matrix method to calculate the scattering
coefficients of the outgoing scattered wave.  The T-matrix describes
the scattering process by relating the scattering wave to the incident
wave:
\begin{align}
	\begin{pmatrix}
		p_{mn} \\
		q_{mn}
	\end{pmatrix}
	= \textbf{T}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}
MSTM computes the T-matrix of the dimer, and then passes the T-matrix
to \textit{ott} to compute the optical forces. The benefit of this
approach is that MSTM can accurately compute the T-matrix for any
arbitrary collection of spheres once, and then \textit{ott} can
calculate the optical forces and torques. The total displacement is
the sum of the optical and Brownian contributions:
\begin{align}
  \langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle
  &= 2D_{ij} \Delta t
  \\
  \Delta q^{opt} &= \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F}
  \\
\Delta q^{tot} &= \Delta q^{\textbf{B}} + \Delta q^{opt}
\end{align}
where \textbf{F} is the a 6-component vector that describes the
optical force and torque components.  Brownian OT returns the dimer's
motion as a \textit{csv} file containing for each timestep the
$[x,y,z]$ position of the dimer's centre of mass relative to the trap
focus, and the orientation represented as a quaternion.

@@

Equation~\eqref{eq:conditional} gives the probability that a signal
($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference
orientation $\hat{\rm n}_\alpha$.
%
Using Bayes' theorem, the probability that the dimer is in reference
orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
$y_3$) is measured, can be estimated:
%.
\begin{align}
  \label{eq:bayes}
  p(\hat{\bf n}_\alpha| y_1,y_2,y_3)
  &=
    \frac{p(y_1,y_2,y_3|\hat{\bf n}_\alpha)
    p(\hat{\bf n}_\alpha)}{p(y_1, y_2, y_3)}
\end{align}
where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior
estimates of the distributions of particle orientations and
instantaneous signals, respectively.
%
Without any prior evidence we assume that the orientation priori of our dimer $p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$ covered in Section~\ref{sec:test}.  The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).  This is given by taking the discrete integral over collection of reference orientations.
\begin{align}
  p(y_1, y_2, y_3)
  =
  \sum_{\alpha=1}^{n_{\rm ref}}
  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
  p(\hat{\bf n}_\alpha)
\end{align}
From \eqref{eq:bayes} we get a mass probability distribution that denotes the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a signal ($y_1$, $y_2$, $y_3$). In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).

@@

Using our simulation from Section~\ref{sec:brownian} we calculated the instantaneous signal from our dimer at each time step using MSTM.  We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation. To visualise our model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time. For comparison, we also plotted the radial distance between the closest reference orientation, denoted $\hat{\bf n}_{best}$, and the dimer's instantaneous orientation. The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another. If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in \ref{fig:uniform}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/uniform_pn.png}
	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.}
	\label{fig:uniform}
\end{figure}

The reasoning for this lies with the fact that differences in reference signals and reference orientation spacing are not directly correlated with each other. In some cases, two reference orientation will have identical signal responses. To remedy this we need to account for the fact that a trapped dimer's motion is severely limited due to the trap stiffness. Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:

@@

where $\beta$ is a weighting factor that controls the significance of our priori estimate. The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}. By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation. To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

The Brownian OT software was used to simulate the motion of a trapped
dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
$10$\,seconds of entering the optical trap.  The initial orientation
was assumed as strictly vertical (in line with the beam propagation
direction). The dimer's position orientation was recorded every
$10\,{\rm \mu s}$ for using as a test dataset for our model.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/traj.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/pos.png}
	\end{subfigure}
	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a
full $180^{\circ}$ rotation upon entering the trap.  Typically
horizontal alignment of a dimer is unstable and will result in the
particle rotating to align along its vertical axis.  It is interesting
to note that the dimer is furthest from the trap centre as it goes
into a horizontal orientation before drawing closer again as it
inverts completely. Further simulations of differently sized dimers
showed similar results, but only when $a_1 \geq 2a_2$.  Dimers with
more symmetrical size ratios immediately aligned into a fixed vertical
position.
%
In the simulations of Vigilante~\emph{et~al}.\
\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
investigated; their findings showed that the optical torque on the
dimer goes to zero while aligned vertically and is at its maximum in a
horizontal alignment.  Therefore, the inversion of an asymmetric dimer
suggests that if the size difference is significant the optical torque
goes is minimal for a dimer in both horizontal and vertical
orientations. Once we have our experimental set up complete we can
confirm this by trapping an asymmetric dimer and monitoring its
orientation.  The data from \figurename{ 2a} was used to test the
efficacy of our predictive model.  To visualise how well our model
tracks the dimer's orientation we plot the radial distance between our
model's prediction and the true orientation, and between the best
possible orientation and the true orientation.  For a perfect
estimation these two plots should overlap, while we cannot achieve an
exact estimation we can optimise it by adjusting the expected signal
error and weighting factor.

@@

A key assumption of Eq.~\eqref{eq:conditional} is that we assume our
detected light signal will inherently have some degree of noise. This
is the central premise that allows us to form
Eq.~\eqref{eq:conditional}, understanding the limits of acceptable
signal error is necessary for implementing this to experimental work.
To this end we ran our model while varying the value of $\epsilon$.
\figurename{ 3} shows the performance of the method with a selection
of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
to $1$:

@@

	\centering
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_at_5.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_at_15.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_at_25.png}
	\end{subfigure}
	\caption{Model prediction for signal error of a) $5\%$, b) $15\%$, and c) $25\%$. Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}}
	\label{fig:epsilon}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, there is a
sweet spot for the signal error; with a $5$\% error giving an
inaccurate estimation, and with a $25$\% error giving an alright
estimation but with sudden shifts away from the ideal result. The
reason why our low error estimation performed so poorly is that the
minimum difference between our values of $y_k$ is larger than our
signal error. As such the model could not distinguish between
different reference orientations and was essentially just picking
options at random. This minimum difference determined by our choice of
angles and sets a lower limit of how accurate our signals can be. In
the future we hope to look at increasing the number of reference
orientations to see what the true lower bound is.

The latter issue of a high signal error can be addressed somewhat byadjusting our value of $\beta$ to flatten our probability distribution. We repeated our model estimation at $25\%$ at different values of $\beta$ to see the upper limit we could improve our result to.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.33\textwidth}
		\includegraphics[width=\textwidth]{./Images/Beta_2.png}
	\end{subfigure}
	\begin{subfigure}{0.33\textwidth}
		\includegraphics[width=\textwidth]{./Images/Beta_3.png}
	\end{subfigure}
	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
	\label{fig:beta}
\end{figure}

@@

We wanted to see if our choice of angles plays a significant role in
our model's efficacy. We employed ultranest to analyse the parameter
space for our three angles by sampling across the priori of
$10^{\circ} \ \& \ 180^{\circ}$.   We kept both $\beta$ and $\sigma$
constant during this so we could as these are parameters that can be
changed more freely.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/corneranglesfreed-1.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
	\label{fig:corner}
\end{figure}

Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer. Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the
choice of detection angle. While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we cannot achieve any greater accuracy than dividing the orientation space into $n_{\rm ref}=30$ reference orientations.  We hope to develop our method further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.

As covered in Section~\ref{sec:epsilon} our choice for beta can help to improve our estimation by restricting it in the physical space. In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.

This result only indicates that orientation is not a characteristic that is angularly biased in its scattering. It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}

We have developed a method for interpreting the dynamics of a trapped
particle based purely on the light scattering pattern. In theory the
model can be applied to any characteristic that impacts the light
scattering pattern produced by the trapped entity. The MSTM package is
flexible in calculating the light scattering for a variety of
micro-particles; so long as there is a distinguishable difference in
the light scattering pattern we can apply this method to characterise
the particle to some degree of accuracy. This has potential to be
applied to characterise particle size and shape that are changing
while in the optical trap. The authors would like to acknowledge and
thank the support for this research from the funding provided by the
Leverhulme Trust.\\

@@

\newpage
\appendix
\onecolumn
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appendix}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\label{sec:App1}
\begin{table}[h]
\begin{center}
\caption{\label{tab:A1}
%
Reference Orientations vector components$^*$}
\begin{tabular}{|c|c|c|c|}
\hline\hline
Reference Orientation & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
\hline
$1$  & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$2$  & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$3$  & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$4$  & $ 1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$5$  & $ 0.00000$ &  $ 1.00000$ & $ 0.00000$ \\
$6$  & $ 0.00000$ &  $ 0.00000$ & $ 1.00000$ \\
$7$  & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$8$  & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$9$  & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$10$ & $ 0.00000$ &  $ 0.00000$ & $-1.00000$ \\
$11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$12$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$14$ & $ 0.00000$ &  $ -1.0000$ & $ 0.00000$ \\
$15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
$16$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
$17$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
$18$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$19$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$21$ & $-1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$23$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$24$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$25$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$26$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$27$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$28$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
$28$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
$30$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
\hline\hline
\multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
		\end{tabular}
	\end{center}
\end{table}

>>>>

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/fig2.png}
	\caption{Example dimer in orientation $\bar{\bf s}$}
	\label{fig:dimer}
\end{figure}

With any trapped particle there are two main sources of motion,
Brownian and optical forces.  The dimer's Brownian motion
($\Delta q^{\textbf{B}}$) is simulated by generating a set of
normally-distributed numbers whose covariance is given by the dimer's
diffusion tensor $\textbf{D}$.  The diffusion tensor is computed based
on Farsund's work \cite{Farsund1996}. For computing the optical force
Brownian OT uses the T-Matrix method to calculate the scattering
coefficients of the outgoing scattered wave.  The T-matrix describes
the scattering process by relating the scattering wave to the incident
wave:
\begin{align}
	\begin{pmatrix}
		p_{mn} \\
		q_{mn}
	\end{pmatrix}
	= \textbf{T}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}
MSTM computes the T-matrix of the dimer, and then passes the T-matrix
to \textit{ott} to compute the optical forces. The benefit of this
approach is that MSTM can accurately compute the T-matrix for any
arbitrary collection of spheres once, and then \textit{ott} can
calculate the optical forces and torques. The total displacement is
the sum of the optical and Brownian contributions:
\begin{align}
  \langle \Delta q_i^{\textbf{B}} \Delta q_j^{\textbf{B}}\rangle
  &= 2D_{ij} \Delta t
  \\
  \Delta q^{opt} &= \frac{\Delta t}{k_BT}\textbf{D} \cdot \textbf{F}
  \\
\Delta q^{tot} &= \Delta q^{\textbf{B}} + \Delta q^{opt}
\end{align}
where \textbf{F} is the a 6-component vector that describes the
optical force and torque components.  Brownian OT returns the dimer's
motion as a \textit{csv} file containing for each time step the
$[x,y,z]$ position of the dimer's centre of mass relative to the trap
focus, and the orientation represented as a quaternion.

@@

Equation~\eqref{eq:conditional} gives the probability that a signal
($y_1$, $y_2$, $y_3$) will be measured if a dimer is in reference
orientation $\hat{\rm n}_\alpha$.
%
Using Bayes' theorem, the probability that the dimer is in reference
orientation $\hat{\bf n}_\alpha$, given that the signal ($y_1$, $y_2$,
$y_3$) is measured, can be estimated:
%.
\begin{align}
  \label{eq:bayes}
  p(\hat{\bf n}_\alpha| y_1,y_2,y_3)
  &=
    \frac{p(y_1,y_2,y_3|\hat{\bf n}_\alpha)
    p(\hat{\bf n}_\alpha)}{p(y_1, y_2, y_3)}
\end{align}
where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior
estimates of the distributions of particle orientations and
instantaneous signals, respectively.
%
Without any prior evidence we assume that the orientation priori of our dimer $p(\hat{\bf n}_{\alpha})$ is uniform. Knowledge of the dimer from previous measurement can be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$, covered in Section~\ref{sec:test}.  The latter prior $p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$).  This is given by taking the discrete integral over collection of reference orientations.
\begin{align}
  p(y_1, y_2, y_3)
  =
  \sum_{\alpha=1}^{n_{\rm ref}}
  p(y_1, y_2, y_3|\hat{\bf n}_\alpha)
  p(\hat{\bf n}_\alpha)
\end{align}
From \eqref{eq:bayes} we get a mass probability distribution that denotes the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ given a signal ($y_1$, $y_2$, $y_3$). In order to evaluate the accuracy of our method, we tested it using a Brownian dynamics simulation of the dimer (see Section~\ref{sec:brownian}).

@@

Using our simulation from Section~\ref{sec:brownian} we calculated the instantaneous signal from our dimer at each time step using MSTM.  We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation. To visualise our model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time. For comparison, we also plotted the radial distance between the closest reference orientation, denoted $\hat{\bf n}_{best}$, and the dimer's instantaneous orientation. The dotted line indicates the maximum radian distance ($0.896$ radians) between two neighbouring reference orientations, if we are under this line then we know our estimate and the best result are beside one another. If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is poor, as seen in Figure~\ref{fig:uniform}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/fig3.png}
	\caption{Model's estimation of dimer orientation while assuming $p(\hat{\bf n}_\alpha)$ to be uniform.}
	\label{fig:uniform}
\end{figure}

The reasoning for this lies with the fact that differences in reference signals and reference orientation spacing are not directly correlated with each other. In some cases, two reference orientation will have identical signal responses. To remedy this we need to account for the fact that a trapped dimer's motion i limited due to the trap stiffness. Therefore, the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was defined at each time step by accounting for the physical distance between our previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha(t)$:

@@

where $\beta$ is a weighting factor that controls the significance of our priori estimate. The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}. By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation. To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:

\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

The Brownian OT software was used to simulate the motion of a trappeddimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
$10$ seconds of entering the optical trap.  The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10\,{\rm \mu s}$ for using as a test dataset for our model.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig4a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig4b.png}
	\end{subfigure}
	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed sa similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.  Dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment.  Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque goes is minimal for a dimer in both horizontal and vertical orientations. Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.  The data from Figure~\ref{fig:motion} was used to test the efficacy of our predictive model.  To visualise how well our model tracks the dimer's orientation we plot the radial distance between our model's prediction and the true orientation, and between the best possible orientation and the true orientation.  For a perfect estimation these two plots should overlap, while we cannot achieve an exact estimation we can optimise it by adjusting the expected signal error and weighting factor.

@@

A key assumption of Eq.~\eqref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise. This is the central premise that allows us to form Eq.~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work. To this end we ran our model while varying the value of $\epsilon$, Figure~\ref{fig:epsilon} shows the performance of the method with a selection of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set to $1$:

@@

	\centering
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/fig5a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/fig5b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/fig5c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of a) $5\%$, b) $15\%$, and c) $25\%$. Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}}
	\label{fig:epsilon}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, there is a sweet spot for the signal error; with a $5$\% error giving an inaccurate estimation, and with a $25$\% error providing a satisfactory estimate but with sudden shifts away from the ideal result. A reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error. As such the model could not distinguish between different reference orientations and was essentially just picking options at random. This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be. In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.

The latter issue of a high signal error can be addressed somewhat by adjusting our value of $\beta$ to flatten our probability distribution. We repeated our model estimation at $25\%$ with different values of $\beta$ to see the upper limit we could improve our result to.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.33\textwidth}
		\includegraphics[width=\textwidth]{./Images/fig6a.png}
	\end{subfigure}
	\begin{subfigure}{0.33\textwidth}
		\includegraphics[width=\textwidth]{./Images/fig6b.png}
	\end{subfigure}
	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
	\label{fig:beta}
\end{figure}

@@

We wanted to see if our choice of angles plays a significant role in our model's efficacy. We employed UltraNest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$.  We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/fig7.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
	\label{fig:corner}
\end{figure}

Figure~\ref{fig:corner} is the resulting corner plot from our UltraNest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer. Overall from the corner plot provided, it is clear that there is no direct correlation between our method's estimation of orientation and the choice of detection angle. This result only indicates that orientation is not a characteristic that is angularly biased in its scattering. It might well be that for other characteristics, such as size or shape, the choice of detection angles may have a more direct impact on our model's performance.

While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we cannot achieve any greater accuracy than dividing the orientation space into $n_{\rm ref}=30$ reference orientations.  We hope to develop our method further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.

As covered in Section~\ref{sec:epsilon} our choice for beta can help to improve our estimation by restricting it in the physical space. In addition depending on our choice of angles we are also setting the lower limit for our signal error $\epsilon$, so it may be preferable to choose a selection of angles that allows for a large signal error and then adjust the weighting factor to improve upon our estimation.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}

We have developed a method for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to characterise particle size and shape that are changing while in the optical trap. We hope to in the future apply this to our own experimental work by charactering the orientation of a trapped asymmetric particle.\\

@@

\newpage
\appendix
\onecolumn
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Appendix}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\begin{table}[h]
\begin{center}
\caption{\label{tab:A1}
%
Reference Orientations vector components$^*$}
\begin{tabular}{|c|c|c|c|}
\hline\hline
Reference Orientation & $\hat{\bf n}_{\alpha,\ x}$ &  $\hat{\bf n}_{\alpha,\ y}$ &  $\hat{\bf n}_{\alpha,\ z}$ \\
\hline
$1$  & $ 0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$2$  & $ 0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$3$  & $ 0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$4$  & $ 1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$5$  & $ 0.00000$ &  $ 1.00000$ & $ 0.00000$ \\
$6$  & $ 0.00000$ &  $ 0.00000$ & $ 1.00000$ \\
$7$  & $ 0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$8$  & $ 0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$9$  & $ 0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$10$ & $ 0.00000$ &  $ 0.00000$ & $-1.00000$ \\
$11$ & $ 0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$12$ & $ 0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$13$ & $ 0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$14$ & $ 0.00000$ &  $ -1.0000$ & $ 0.00000$ \\
$15$ & $ 0.29588$ &  $-0.29588$ & $-0.90825$ \\
$16$ & $ 0.90825$ &  $-0.29588$ & $-0.29588$ \\
$17$ & $ 0.29588$ &  $-0.90825$ & $-0.29588$ \\
$18$ & $-0.29588$ &  $ 0.29588$ & $ 0.90825$ \\
$19$ & $-0.90825$ &  $ 0.29588$ & $ 0.29588$ \\
$20$ & $-0.29588$ &  $ 0.90825$ & $ 0.29588$ \\
$21$ & $-1.00000$ &  $ 0.00000$ & $ 0.00000$ \\
$22$ & $-0.29588$ &  $ 0.29588$ & $-0.90825$ \\
$23$ & $-0.90825$ &  $ 0.29588$ & $-0.29588$ \\
$24$ & $-0.29588$ &  $ 0.90825$ & $-0.29588$ \\
$25$ & $-0.29588$ &  $-0.29588$ & $ 0.90825$ \\
$26$ & $-0.90825$ &  $-0.29588$ & $ 0.29588$ \\
$27$ & $-0.29588$ &  $-0.90825$ & $ 0.29588$ \\
$28$ & $-0.29588$ &  $-0.29588$ & $-0.90825$ \\
$28$ & $-0.90825$ &  $-0.29588$ & $-0.29588$ \\
$30$ & $-0.29588$ &  $-0.90825$ & $-0.29588$ \\
\hline\hline
\multicolumn{4}{l}{\small *Orientation vector points from centre of sphere 1 to centre of sphere 2.} \\
		\end{tabular}
	\end{center}
\label{sec:App1}
\end{table}
